# Exercise 3


## 1. Getting and reading the .csv file

```{r, include=T}
getwd()
alc <- read.csv(file = "/Users/streetman/IODS-project/data/alc.csv", header = T, sep=",")
```


```{r, include = T}
library(dplyr)
colnames(alc)
```
#### Commentary
The dataset includes the variables above indicating the questions to students relating to their alcohol usage



## 2. Relationships between high and low alcohol consumption, and choosing the parameters for this exploration

### Choosing the parameters

```{r, include=T}
library(tidyr)
glimpse(alc)
```

#### Commentary
These four variables were chosen:
1. Age (numeric)
2. Famsize (binary: less or equal to 3, or greater than 3)
3. Failures (numeric: 0-3, or 4 if smth else)
4. Absences (numeric: 0 to 93)

Hypothesis: there is a correlation between the high alcohol consumption and the chosen variables when analyzed in a univariate method to observe correlation
H0: there is no correlation between a chosen variable and high alcohol consumption
H1: there is significant correlation between a chosen variable and high alcohol consumption



## 3. Distributions between the chosen variables above and the relations to alcohol consumption

### Numerical overview of data
```{r, include=T}
summary(alc)
```
#### Commentary
For example in variable "absences" the mean absence value is 4.5 and maximum is 45. Range is wide, although 3rd quartile is 6.0, therefore most are within values 0 to 6.0.



### Graphical overview of data

### Distribution of the chosen variables
```{r, include=T}
library(tidyr); library(dplyr); library(ggplot2)
alc_selected <- select(alc, one_of(c("age", "famsize", "failures", "absences")))
gather(alc_selected) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

#### Commentary
For example in variable "age" we can detect that mainly the participants are below 19 years old.



### Variables vs. alcohol usage by histograms

### Age

```{r, include=T}
g1 <- ggplot(alc, aes(x = high_use, y = age))
g1 + geom_boxplot(aes(col=sex)) + ylab("age") + ggtitle("Students alcohol consumption and age")
```

#### Commentary
Those with high alcohol consumption were mainly male and their age range is wider compared to female within this consumption category.



### Famsize
```{r, include=T}
g1 <- ggplot(data = alc, aes(x = alc_use, fill = famsize))
g1 + geom_bar()
```

#### Commentary
It seems that more higher the alcohol consumption then more equal are the family sizes. With lower alcohol consumption the family sizes were considerable larger.


### Failures
```{r, include=T}
alc$failures <-  as.factor(alc$failures)
g1 <- ggplot(alc, aes(x = high_use, fill = failures))
g1 + geom_bar()
```

#### Commentary
Somewhat surpringly, in lower alcohol consumption the amount of failures is higher.


### Absences
```{r, include=T}
g1 <- ggplot(alc, aes(x = high_use, y = absences))
g1 + geom_boxplot(aes(col=sex)) + ylab("absences") + ggtitle("Students alcohol consumption and absences")
```

#### Commentary
It seems that more the absences then more are is the high alcohol consumption with both sexes.



## 4. Logistic regression model of the chosen variables

### Fitting the LR model

```{r, include = T}
## Fit a linear model
alc$failures <-  as.numeric(alc$failures)
alc$famsize <-  as.numeric(alc$famsize)
model_fit <- lm(high_use ~ age + famsize + failures + absences, data = alc)

## Print out a summary of the model
summary (model_fit)
```

```{r, include = T}
## Fit a linear model
alc$failures <-  as.numeric(alc$failures)
model_fit <- lm(high_use ~ age + failures + absences, data = alc)

## Print out a summary of the model
summary (model_fit)
```

```{r, include = T}
## Fit a linear model
alc$failures <-  as.numeric(alc$failures)
alc$famsize <-  as.numeric(alc$famsize)
model_fit <- lm(high_use ~ failures + absences, data = alc)

## Print out a summary of the model
summary (model_fit)
```
#### Commentary
The target variable is high_use (of alcohol). For the chosen explanatory variables age, family size, failures and absences: with age and family size there is no statistically significant relationship with the target variable (p=0.089 and 0.170), therefore family size was removed.

After that, with variable age, there is no statistically significant relationship with the target variable (p=0.089), therefore it was removed.

After that, with variables failures and absences, there is statistically significant correlation (p=0.004 and <0.000) to high_use (of alcohol) with an explanatory value of 0.07 (mult. R-squared) which means that these variables explains approx. 7% of the high_use (of alcohol) among the participants.





### Odds rations and CI

```{r, include=T}
### Fitting the logistic regressio model
m <- glm(high_use ~ failures + absences, data = alc, family = "binomial")

### Compute odds ratios (OR)
OR <- coef(m) %>% exp

### Compute confidence intervals (CI)
CI <- confint(m) %>% exp

### Printing the output
cbind(OR, CI)
```
#### Commentary
Both variables', failures and absences, ORs indicate higher odds for high_use (of alcohol) within these variables; yet, the CI crosses value 1 with the variable failures, therefore it seems to inadequate for showing statistical significance in this sense. By this analysis, from the chosen variables, absenses seems to be the one with most statistical significance in its relation to high_use (of alcohol).



## 5. Predictive power of LR model

```{r, include=T}
## Fit the model
m <- glm(high_use ~ absences, data = alc, family = "binomial")

## Predict() the probability of high_use
probabilities <- predict(m, type = "response")

## Add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

## Use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = probability > 0.5)

## Tabulate the target variable versus the predictions
results <- table(high_use = alc$high_use, prediction = alc$prediction)

results

tab <- table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table() %>% addmargins()

tab
```

```{r, include=T}
## Total proportion of inaccurately classified individuals
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

loss_func(class = alc$high_use, prob = alc$probability)
```
#### Commentary
After LR analysis, it was stated that variable "absences" has the most statistical significance it relation to high_use (of alcohol), therefore chosen here to test the predictive power of the LR model.

After calculation, the value of 0.2879 indicates that approx. 29% are incorrect predictions with this model, yet 71% are correct ones, therefore this LR model is useful.
