# Exercise 2


## 1. Reading the data and exploring the structure

```{r, include = T}
students2014 <- read.csv(file = "/Users/streetman/IODS-project/data/lrn14.csv", header = T, sep=",")
students2014 <- students2014[,-1]
str(students2014)
dim(students2014)
```

### Commentary
The data is a subset from the main data, which includes different questions to students and also some demographical variables for the purpose of doing a survey of approaches to learning. This subdata consists 166 rows as observations and 7 columns as variables: gender, Age, Attitude, deep (deep learning related questions), stra (strategic learning related questions), surf (surface learning related questions).



## 2. Graphical overview
```{r, include = T}
library(GGally)
ggpairs(students2014, mapping = aes(col=gender, alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
```

### Commentary
Based on visualization we can detect for example that respondends were almost twice as much females (N=110) and their age varied between 17 and 55, and mean age was 25. Regarding correlation the strongest is between attitude and points and lowest between deep and points.



## 3. Regression model and fitting variables

```{r, include = T}
## Fit a linear model
model_fit <- lm(Points ~ Attitude + stra + surf, data = students2014)

## Print out a summary of the model
summary (model_fit)
```

```{r, include = T}
## Fit a linear model
model_fit <- lm(Points ~ Attitude + stra, data = students2014)

## Print out a summary of the model
summary (model_fit)
```

```{r, include = T}
## Fit a linear model
model_fit <- lm(Points ~ Attitude, data = students2014)

## Print out a summary of the model
summary (model_fit)
```


### Commentary and interpretation
The target variable is Points. For the chosen explanatory variables Attitude, stra and surf: with stra and surf there is no statistically significant relationship with the target variable (p=0.117 and 0.466), therefore surf was removed.

After that, with variable stra, there is no statistically significant relationship with the target variable (p=0.089), therefore it was removed.

After that, with variable attitude, there is statistically significant correlation (p = <0.000) to points with a explanatory value of 0.1856 (adj. R-square) which mean that this variable explains approx. 19% of the points achieved by students.



## 4. Diagnostic plots (Residuals vs. Fitted values, Normal Q-Q plot and Residuals vs Leverage)

```{r, include = T}
my_model2 <- lm(Points ~ Attitude, data = students2014)

plot(my_model2, which = c(1,2,5))
```

### Commentary
The residuals are the differences between the predicted and observed value in a given point. "Residuals vs. Fitted": assumption that errors don't depend on variable attitude is valid since the plots are evenly spread without any specific pattern. "Normal Q-Q": assumption is that the residuals are normally distributed and this supports it. "Residuals vs. Leverage": in general observations are not having unusually high impact.



