# Exercise 6

## 1. Dataset "RATS"

### Loading and exploring

```{r, include=T}

library(ggplot2)
library(tidyr)
library(dplyr)

ratsl <- read.csv("/Users/streetman/IODS-project/data/RATSL.csv", row.names = 1)

ratsl$Group <- factor(ratsl$Group)
ratsl$ID <- factor(ratsl$ID)

ratsl <- ratsl %>% group_by(Time) %>% mutate(stweight = (Weight - mean(Weight)) / sd(Weight)) %>% ungroup()

str(ratsl)

```

#### Commentary

16 rats, divided into 3 groups based on different diets


### Graphics

```{r, include=T}

ggplot(ratsl, aes(x = Time, y = Weight, linetype = ID)) + geom_line() + facet_grid(. ~ Group, labeller = label_both) + scale_linetype_manual(values = rep(1:6, times=4)) + theme(legend.position = "none") +   scale_y_continuous(limits = c(min(ratsl$Weight), max(ratsl$Weight)))

```

#### Commentary

Rats are gaining weight with the difference in group 1 that they are lighter in all points with a variance also between the subject with a decrease of value.

### Summarizing the dataset

```{r, include=T}

n <- ratsl$Time %>% unique() %>% length()

ratssum <- ratsl %>% group_by(Group, Time) %>% summarise(mean = mean(Weight), se = ( sd(Weight) / sqrt(n) )) %>% ungroup()

glimpse(ratssum)

par(mfrow = c(1,2))

ggplot(ratssum, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  theme(legend.position = c(0.9,0.5)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

ggplot(ratssum, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.9,0.5)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

```

#### Commentary

In time, means are increasing in both groups.
 

### Box-plotting
 
```{r, include=T}

ggplot(ratsl, aes(x = factor(Time), y = Weight, fill = Group)) + geom_boxplot()

```
 
 
### Elements for summary (visualizing, excluding)
 
```{r, include=T}

rat8s <- ratsl %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

glimpse(rat8s)

ggplot(rat8s, aes(x=Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white")

```

 
```{r, include=T}

glimpse(rat8s)

rat8s1 <- filter(rat8s, Group == 2 & mean < 550 | Group == 1 & mean > 260 | Group == 3 & mean > 500)

glimpse(rat8s1)

ggplot(rat8s1, aes(x=Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white")

```

 
### Comparing means
 
```{r}

summary(aov(mean ~ Group, data = rat8s1))

```
 
#### Commentary

Using T-test, outliers excluded, no difference between the groups yet no signicance either
 
 
```{r, include=T}

kruskal.test(mean ~ Group, data = rat8s1)

```

#### Commentary

With this test we can find signicance

### More analysis (also adding from the original dataset)
 
```{r, include=T}

rats <- read.csv("/Users/streetman/IODS-project/data/rats.csv", row.names = 1)

rat8s2 <- rat8s %>% mutate(baseline = rats$WD1)

glimpse(rat8s2)


fit <- lm(mean ~ baseline + Group, data = rat8s2)
summary(fit)

```

#### Commentary

Baseline is showing relatedness to means in later phase, yet, after controlment the baseline no signicance between the group means.


## 2. Dataset "BPRS"


### Loading
	 
```{r, include=T}

bprsl <- read.csv(file = "/Users/streetman/IODS-project/data/bprsl.csv", row.names = 1)

bprsl$treatment <- factor(bprsl$treatment)

bprsl$subject <- factor(bprsl$subject)

glimpse(bprsl)

```
	 

### Graphical

```{r, include=T}
ggplot(bprsl, aes(x = weeks, y = bprs, group = subject)) + geom_line() + facet_grid(. ~ treatment, labeller = label_both) 

```

#### Commentary
Points are decreasing in time with all individuals.
	 

### Visualizing

```{r, include=T}

bps_reg <- lm(bprs ~ weeks + treatment, data=bprsl)
summary(bps_reg)

```
	 
#### Commentary

LRM of BPRS with explanatory variables of response, weeks and treatment group indicates that decreasing in time with the latter variable being not significant.

	 
### RIM (Random Intercept Model)

```{r, include=T}

library(lme4)

bps_ref <- lmer(bprs ~ weeks + treatment + (1|subject), data = bprsl, REML=FALSE)

summary(bps_ref)
	
```
	 
#### Commentary

Within-subject elements are not included.


### RIM and RSM (Random Slope Model)
	 
```{r, inlclude=T}

bps_ref2 <- lmer(bprs ~ weeks + treatment + (weeks|subject), data = bprsl, REML=FALSE)

summary(bps_ref2)

```

### Analyzing with ANOVA

```{r, include=T}

anova(bps_ref, bps_ref2)

```
	 
#### Commentary

Model 2 is better


### RIM and SLM (interaction included)

```{r, include=T}

bps_ref3 <- lmer(bprs ~ weeks * treatment + (weeks|subject), data = bprsl, REML=FALSE)

summary(bps_ref3)

```

#### Commentary

The model indicates that the BPRS slopes with a decreasing element are higher for individuals in treatment group 2 in comparison.

```{r, include=T}

anova(bps_ref2, bps_ref3)

```
	 
#### Commentary

Model 3 doesn't show any superiority, therefore, not better


### Fitting and plotting
	 
```{r, include=T}

Fitted <- fitted(bps_ref2)

bprsl <- bprsl %>% mutate(Fitted)

p1 <- ggplot(bprsl, aes(x = weeks, y = bprs, group = subject)) + geom_line() + facet_grid(. ~ treatment, labeller = label_both) + scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 1)) +  ylim(0,100) + theme(legend.position = "top") + ggtitle("Observed")

p2 <- ggplot(bprsl, aes(x = weeks, y = Fitted, group = subject)) +
	  geom_line() + facet_grid(. ~ treatment, labeller = label_both) +
	  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 1)) +
	  ylim(0, 100) +
	  theme(legend.position = "top") + ggtitle("Fitted")
	
p1

p2

```

