# Exercise 5

```{r, include=T}
library(data.table)
library(GGally)
library(corrplot)
library(tidyr)
library(ggplot2)
```


## 1. Getting and reading the .csv file

```{r, include=T}
getwd()
human <- read.csv(file = "/Users/streetman/IODS-project/data/human.csv", header = T, sep=",")
dim(human)
str(human)
summary(human)

```

### Commentary
The dataset is about United Nations Development Programme. Variables are in respective order:

Ratio of secondary education of female/male (Edu2.FM)
Ratio of labour forced female/male (Labo.FM)
Expected years of schooling (Edu.Exp)
Life expectancy at birth (Life.Exp)
Gross National Income per capita (GNI)
Maternal mortality ratio (Mat.Mor)
Adolescent birth rate (Ado.Birth)
Percentage of female representatives in parliament (Parli.F)

We can detect for example expected years of education is 5.40 years at minimum with a median of 13.18 years. Regarding life expectancy, the minimum value is 49 years of age and mean value being approx. 72 years.


## 2. Graphical overview


```{r, message=FALSE, warning=FALSE, fig.width=14, fig.height=18}

human$GNI <- as.integer(human$GNI)

human_ <- human[,-1]

dim(human_)
str(human_)

ggpairs(human_)

```

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}

cor(human_) %>% corrplot(method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex = 0.6)

```

### Commentary

Graphical illustration demonstrates distributions of the data variables and their correlation to each other.

Correlation matrix is giving further supportive evidence of the correlation between data parameters.


## 3. PCA

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=12}
pca_human <- prcomp(human_)

s <- summary(pca_human)

pca_pr <- round(100*s$importance[2, ], digits = 1)

pca_pr
```

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=12}

biplot(pca_human, choices = 1:2, cex=c(0.8, 1), col=c("grey40", "deeppink2"))

```


### Commentary

The first table is indicating that majority of the data is explained by principal component 1 (PC1) so the data could be summarized towards that dimension.

The biplot graph indicates that the PC1 explains majority of cross-observation variation with maternal mortality as a driving component.


## 4. Standardize
```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=12}

human_std <- scale(human_)
summary(human_std)
pca_human <- prcomp(human_std)

s <- summary(pca_human)
pca_pr <- round(100*s$importance[2, ], digits = 1)
pca_pr
```


```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=12}
biplot(pca_human, choices = 1:2, cex=c(0.8, 1), col=c("grey40", "deeppink2"))
```

### Commentary

GNI, Edu2.Ratio, Edu.Exp and Life.Exp have negative weights on PC1.

Mat.Mor and Ado.Birth have positive weights on PC1.

Labo.FM has the most positive effect on PC2.


## 5. Interpretation
Components related to e.g. women education and life expectancy seem to explain differences between observations due to their largest weights on PC1.

Component related to women labour seems to affect the PC2 the most.


## 6. Loading the Tea dataset


```{r, message=FALSE, warning=FALSE}
library(FactoMineR)

data(tea)

str(tea)
dim(tea)
```


```{r, message=FALSE, warning=FALSE}
library(dplyr)

keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- dplyr::select(tea, one_of(keep_columns))

# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x =  element_text(angle = 45, hjust = 1, size = 8))
```

### MCA

```{r, include=T}
mca <- MCA(tea_time, graph = FALSE)

summary(mca)

plot(mca, invisible=c("ind"), habillage = "quali")
```

### Commentary

The graph shows associations between the categories. Dimensions are explanatory percentages of the variation with respective values shown in prackets.

Dimension 1 and Dimension 2 explain 18,4 % and 17,1% of the variation, respectively.


