# Exercise 3


## 1. Getting and reading the .csv file

```{r, include=T}
getwd()
alc <- read.csv(file = "/Users/streetman/IODS-project/data/alc.csv", header = T, sep=",")
```


```{r, include = T}
library(dplyr)
colnames(alc)
```
The dataset includes the variables above indicating the questions to students relating to their alcohol usage



## 2. Relationships between high and low alcohol consumption, and choosing the parameters for this exploration

### Choosing the parameters

```{r, include=T}
library(tidyr)
glimpse(alc)
```

1. Age (numeric)
2. Famsize (binary: less or equal to 3, or greater than 3)
3. Failures (numeric: 0-3, or 4 if smth else)
4. Absences (numeric: 0 to 93)

Hypothesis: there is correlation between the high alcohol consumption and the chosen variables when analyzed in a univariate method to observe correlation
H0: there is no correlation between a chosen variable and high alcohol consumption
H1: there is significant correlation between a chosen variable and high alcohol consumption



## 3. Distributions between the chosen variables above and the relations to alcohol consumption

### Numerical overview of data
```{r, include=T}
summary(alc)
```


### Graphical overview of data

### Distribution of the chosen variables
```{r, include=T}
library(tidyr); library(dplyr); library(ggplot2)
alc_selected <- select(alc, one_of(c("age", "famsize", "failures", "absences")))
gather(alc_selected) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

### Variables vs. alcohol usage by histograms

### Age

```{r, include=T}
g1 <- ggplot(alc, aes(x = high_use, y = age))
g1 + geom_boxplot(aes(col=sex)) + ylab("age") + ggtitle("Students alcohol consumption and age")
```

### Famsize
```{r, include=T}
g1 <- ggplot(data = alc, aes(x = alc_use, fill = famsize))
g1 + geom_bar()
```

### Failures
```{r, include=T}
alc$failures <-  as.factor(alc$failures)
g1 <- ggplot(alc, aes(x = high_use, fill = failures))
g1 + geom_bar()
```

### Absences
```{r, include=T}
g1 <- ggplot(alc, aes(x = high_use, y = absences))
g1 + geom_boxplot(aes(col=sex)) + ylab("age") + ggtitle("Students alcohol consumption and absences")
```



## 4. Logistic regression model of the chosen variables

### Fitting the LR model

```{r, include = T}
## Fit a linear model
model_fit <- lm(high_use ~ age + famsize + failures + absences, data = alc)

## Print out a summary of the model
summary (model_fit)
```

```{r, include = T}
## Fit a linear model
model_fit <- lm(high_use ~ famsize + failures + absences, data = alc)

## Print out a summary of the model
summary (model_fit)
```

```{r, include = T}
## Fit a linear model
model_fit <- lm(high_use ~ failures + absences, data = alc)

## Print out a summary of the model
summary (model_fit)
```


### Odds rations and CI

```{r, include=T}
### Fitting the logistic regressio model
m <- glm(high_use ~ failures + absences, data = alc, family = "binomial")

### Compute odds ratios (OR)
OR <- coef(m) %>% exp

### Compute confidence intervals (CI)
CI <- confint(m) %>% exp

### Printing the output
cbind(OR, CI)
```


## 5. Predictive power of LR model

```{r, include=T}
## Fit the model
m <- glm(high_use ~ failures + absences, data = alc, family = "binomial")

## Predict() the probability of high_use
probabilities <- predict(m, type = "response")

## Add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

## Use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = probability > 0.5)

## Tabulate the target variable versus the predictions
results <- table(high_use = alc$high_use, prediction = alc$prediction)

results

tab <- table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table() %>% addmargins()

tab
```

```{r, include=T}
## Total proportion of inaccurately classified individuals
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

loss_func(class = alc$high_use, prob = alc$probability)
```
